---
title: "lab_08"
author: "Khushboo Rathore"
date: "2023-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. We'll be making some charts, working with dates and retrieving Census data.

```{r loadings}
library(tidyverse)
library(lubridate)
library(tidycensus)
library(dplyr)
```

## Load and modify data

**Task** Create a codeblock and load the following data from the data folder:

* Prince George's County 911 Overdose Calls

You will need to create columns for the date, week and month based on the existing `datetime` column.

```{r}
census_api_key("4e2627391241d2d3419becb0abe8b0a8cffe8aab")
pg_overdose_calls <- read_csv("data/prince_georges_2022_overdoses.csv") 

pg_od_clean <- pg_overdose_calls %>% 
  mutate(date = as.Date(datetime)) %>% 
  mutate(week = week(date)) %>% 
  mutate(month = month(date, label = TRUE))
```


## Questions 

**Q1.** Which month saw the greatest percentage of total calls? Create a dataframe that calculates the percentage of all calls that each month's calls represents. Make a bar chart from that dataframe. Your bar chart must have:

* A clear title that states the main idea/finding
* Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
* Readable bars - the values shouldn't be overlapping

Compare the results here to those from the pre_lab_08 bar chart - are there major differences in the months with the highest and lowest figures? Describe that below.

**A1.** 
December saw the most calls, which is a stark contrast with Baltimore County. The highest months in PG are colder, while those in Baltimore County as in warmer months. Specifically, December is at the top in PG, while it's near the bottom in Baltimore County. While the coldest months (Jan, Feb) are near the bottom on both, June, July and August are all near the bottom in PG when they were at the top in Baltimore County.
```{r}
count_calls = nrow(pg_od_clean)
monthly_percents <- pg_od_clean %>% 
  group_by(month) %>% 
  summarize(
    num_calls = n()
  ) %>% 
  mutate(percent = round(num_calls/count_calls*100,2))

percents_chart <- monthly_percents %>% 
  ggplot() + 
  geom_bar(aes(x = reorder(month,percent), weight = percent)) +
  coord_flip() + 
  labs(
    title = "More 911 Overdose Calls in Cooler Months",
    x = "Month",
    y = "Percentage of all 2022 calls (%)",
    caption = "Source: Prince George's County EMS"
  ) +
  theme_minimal()
percents_chart
```

**Q2.** Let's visualize this data on a weekly basis using a line chart. As in Q1, generate a dataframe with the total number of calls for each week, and then create a line chart to show the distribution of calls over time. Your line chart must have:

* A clear title that states the main idea/finding
* Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
* Readable labels

Describe the pattern of results; you may want to look at the data to dig into particular outliers.

**A2.** 
The low count at the end of the year corresponds with an incomplete 7 day period. The high count explains why July is a bit of an outlier in our monthly percentage data. This may or may not be relevant, but the week of this outlier (week 28, July 09-July 15) was also the week leading up to primary elections. And, it was also the week when the massive storm hit PG county, destroying a lot of the county.

```{r}
weekly_counts <- pg_od_clean %>% 
  group_by(week) %>% 
  summarize(
    num_calls = n()
  )

weekly_chart <- weekly_counts %>% 
  ggplot() +
  geom_line(aes(x = week, y = num_calls)) +
  labs(
    title = "More 911 Overdose Calls as year goes on",
    x = "Week of the Year",
    y = "Number of Calls",
    caption = "Source: Prince George's County EMS"
  )

weekly_chart

## Irrelevant :(
investigato_potato <- pg_od_clean %>% 
  filter(week == 28) %>% 
  group_by(zipcode) %>% 
  summarize(
    calls = n()
  )
```

**Q3.**  A scatterplot is a type of chart that helps us see relationships between two variables. One variable goes on the x axis, the other on the y axis.  For each row/observation in our data, a scatterplot puts a circle (or a "point") where the two variables intersect on a grid. 

Statisticians use scatterplots to show graphically whether one variable is correlated -- related, in a statistical sense -- with another variable.  A classic example is the [relationship between ice cream sales and temperature](https://www.mathsisfun.com/data/scatter-xy-plots.html). The scatterplot below -- press play to load the image -- shows that relationship, that an increase in temperature is associated with an increase in ice cream sales. When it's 12C, sales are 200 dollars, and when it's hotter, 25C, sales are 600 dollars.

```{r}
knitr::include_graphics("https://www.mathsisfun.com/data/images/scatter-ice-cream1.svg")
```

We're going to use a scatterplot a little differently, to get a visual sense of two key variables: 

Our question is: does the median income in a zip code have any relationship to the number of overdose 911 calls in that zip code?

To answer this question, do the following:

1. Generate a dataframe with the number of 911 calls for each zip code.
2. Get data from the Census Bureau showing median household income for Maryland zip codes.
3. Join those two dataframes on their zip code columns, starting with the 911 calls dataframe.
4. Make a scatterplot showing the total calls and median income. I didn't show you how to do this, so look it up! Googling "ggplot scatterplot" is a good start.
5. Give it an appropriate title, source, and x and y axis titles.
6. Add a label for each point that shows the zip code using geom_text() - see some examples of its use at https://ggplot2.tidyverse.org/reference/geom_text.html#ref-examples. Try to make the names as easy to read as possible by avoiding overlap.
7. In the answer space below, describe what you see and answer the questions posed above. In a general sense, what do you think this means? Feel free to consider the actual raw values: how would you report out the main point(s) of this chart?

**A3.**  
Overall, it seems like areas with a lower median income have more 911 overdose calls. The biggest outlier is 20710, which has only 25 calls despite having a very low median income. In all fairness, both 20710 and 20712 have populations of under 10,000 people in comparison to some of the areas with more calls. This leads me to believe that the data and graph may not actually show what I think it does. Rather, we should look at calls per 1000 people versus the median income to get a more accurate picture. I wouldn't necessarily feel comfortable reporting without looking at that data. In the similar vein, 20769 has a population under 10,000 and it mostly older people. So really, I don't think this graph is something I would publish or report on.
```{r}
calls_by_zip <- pg_od_clean %>% 
  group_by(zipcode) %>% 
  summarize(
    count_calls = n()
  )

md_median <- md <- get_acs(geography = "zcta",
              variables = c(medincome = "B19013_001"),
              state = "MD",
              year = 2019)

income_calls <- calls_by_zip %>% 
  inner_join(md_median, by = c("zipcode" = "GEOID"))

call_inc_scat <- income_calls %>% 
  ggplot(aes(x = count_calls, y = estimate,label = zipcode)) +
  geom_point(size = 2, shape = 1, color = "black") + 
  labs(
    title = "Lower Income Areas have the Most 911 Overdose Calls",
    x = "Number of Calls",
    y = "Median Income"
  ) +
  geom_text(check_overlap = TRUE, vjust = "bottom", hjust = "left")

call_inc_scat
```
