---
title: "lab_05"
author: "Khushboo Rathore"
date: "2023-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

- Tabula

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(lubridate)
```

## Get Our PDF

We'll be working with the [911 overdose calls from Baltimore County](https://drive.google.com/file/d/1qkYuojGF_6WKFr5aNQxmewDzcKyOiJFr/view?usp=share_link). You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to extract the tables within it, export that to a CSV file, load it into RStudio and ask some questions.

## Extract Data from PDF Using Tabula

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds or longer.

This PDF has a single table spread over multiple pages to extract. We're going to make a single dataframe from this table, exporting it to a CSV file that you will load into R. In Tabula, highlight the table and click the "Preview & Export Extracted Data" button. You may want to play with including or excluding the column headers - YOU SHOULD HAVE FIVE COLUMNS OF DATA.

Save the CSV (it should be called `tabula-Baltimore County; Carey, Samantha log OD.csv` by default) to your lab_05/data folder.

From there, you will need to read in the data, and add or fix headers if necessary. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)).

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers should not contain spaces and they should have meaningful names, not "x1" or something similar. How you do that is up to you, but you can use select() with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns. Importantly, you'll need to ensure that any columns containing a date actually have a date datatype. Our friend `lubridate` can help with this.

```{r}
baltimore_county_calls <- read_csv("data/tabula-Baltimore County; Carey, Samantha log OD.csv", col_names = c("date", "time", "case_num", "event_type", "location")) %>% 
  mutate(date = mdy(date))

```

## Answer questions

Q1. Write code to generate the number of calls that occurred on each date. Which date in 2022 had the most overdose calls, and how many? Look at the total number of rows in your result and explore the range of dates - based on your result, do you believe there are any days with no overdose calls at all? Explain why or why not.

A1. 
July 14, 2022 and October 4, 2022 both had 23 calls. 
No, I don't. Part of the problem is that the dates in 2022 start on February 6, 2022. This means we do not have data for 36 days of 2022. 365-36 = 329, which is the number of rows that we have. Considering the fact that there is at least one call in all the days we have records for, I find it HIGHLY unlikely that there is a day in our missing set with no calls.
```{r}
bc_calls_date <- baltimore_county_calls %>%
  filter(!year(date) == 2023) %>% 
  group_by(date) %>% 
  count() %>% 
  arrange(desc(n))

```

Q2. You want to understand if there's a pattern in the day of the week that overdose calls are made. Add a column to your dataframe that displays what day of the week each date represents. You should search for how to do that using lubridate. Then write code to calculate the number of calls for each day of the week, and add a column to that result that calculates the percentage of all calls that occurred on each day of the week (so you want a dataframe with the day of the week, total number of calls and the percentage of calls on that day out of the total number of all calls). Describe your findings to me.

A2. 
There isn't a very signifcant difference (in my opinion) of how many overdose calls are made per day of the week. The weekend days (Fri, Sat, Sun) seem to be a little more likely than the middle of the week (Wed, Thurs). Overall, the closer to the weekend, the more overdose calls there are. However, the difference is by less than 3 percentage points.

```{r}
all_calls = nrow(baltimore_county_calls)
bcalls_weekdays <- baltimore_county_calls %>% 
  mutate(weekday = wday(date, label = TRUE)) %>% 
  group_by(weekday) %>% 
  count() %>% 
  mutate(percent = round(n/all_calls*100,2)) %>% 
  arrange(desc(percent))

```

Q3. Now let's look at locations. Which ones have the most calls? How would you describe them (feel free to search for more information on them)? Is there anything about the structure of the original data that might make you less confident in the counts by location or date?

A3. 

The biggest problem in the structure of the data is also reflected in the most common type of location, which is intersections. Dispatchers and people who witness ODs and don't know the address may be getting this information on the fly, leading to less precision. After intersections, Motels and Inns were the second most, then police precincts. Without using string detect to categorize locations, the most common one is "4540 SILVER SPRING RD," which is actually the intersection of Perry Hall Blvd and Silver Spring Road. This indicates another issue. Some addresses may actually be intersections and vice versa. I'm pretty comfortable with the records of the dates, but there are duplicate calls and incidents that are recorded. Without duplicates, there are 3,749 calls. There could be two issues: double records or multiple people calling at once, but getting the time the same to the second seems kind of suspicious.

```{r}
bc_calls_loc <- baltimore_county_calls %>%
  # mutate(location = case_when(
  #   str_detect(location, "PC")~"Police Precinct",
  #   str_detect(location, "INNERDALE")~location,
  #   str_detect(location, "MOTEL| INN")~"Motel/Inn",
  #   str_detect(location, "\\&")~"Intersection",
  #   TRUE~location
  # )) %>% 
  group_by(location) %>% 
  count() %>% 
  arrange(desc(n))

bc_distinct <- baltimore_county_calls %>% 
  distinct(case_num)
```

Q4. What's the best story idea or question you've seen as a result of the work you've done in this lab?

A4. 
I think it would be interesting to look into levels of police presence and response times to locations that have a large volume of calls. In the same vein, I would want to look at how shifts are scheduled for paramedics related to dates and times. Also, I would love to look at socio-economic and racial breakdowns by putting thi data on a map.