---
title: "lab_04"
author: "Khushboo Rathore"
date: "2023-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). Read the questions carefully and attempt to do all of what they ask. Make a list if it helps.

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load three packages for this: the tidyverse, lubridate and janitor.

**Task** load these three packages.

```{r}
# Turn off scientific notation
library(tidyverse)
library(lubridate)
library(janitor)
options(scipen=999)

```

For this lab, we want to investigate spending by Maryland state and local candidates during the 2022 election. For example, we could ask how much money have candidates and committees spent on advertising on Facebook? Which vendors received the most money for media expenses? We have the data, but it's messy - names of payees are spelled differently and capitalized differently - and we need to clean it up. We'll use a combination of RStudio and OpenRefine to do that.

The basic process here is to start in RStudio, export data so that we can use OpenRefine to clean it up, and then bring it back into RStudio to finish our analysis. The function to export data from RStudio is called `write_csv()` and it takes two arguments: the name of the dataframe and the path of the file to write it to, which is in quotes. Like `read_csv()`, but the variable name comes inside. Here's an example, in which you'd replace the dataframe and the path values.

```{r}
write_csv(dataframe, "folder/file.csv")
```


## Load Data

You'll need to load one data set, a CSV file of Maryland campaign expenditures located in the data folder called "maryland_expenses.csv"

**Task** Create a codeblock below, then read the data in and assign it to an appropriate variable name. You'll want to clean up the column names and make sure that any date columns are actually date datatypes.

```{r}
# Load the Maryland expenditures data table, clean the column names and convert columns with dates into actual dates.
maryland_expenses <- read.csv("data/maryland_expenses.csv") %>% 
  clean_names() %>%
  mutate(expenditure_date = mdy(expenditure_date))
# 15 failed to parse, are not dates but instead company names or addresses
```

## Answer questions

**Q1.** You've been assigned to write about Maryland political committees' use of Facebook advertising. There's a category column in the data that has values for different kinds of spending. One of them is "Media". Write code to isolate Media expenses where the name of the payee is not NA. 

Then, using the `write_csv` function, make a CSV file of those media expenses in your data folder. Using OpenRefine, create a project using that CSV file, create a copy of the `payee_name` column called `payee_clean` and then standardize it, focusing your work on the payees that appear most often. 

Pay particular attention to variations of Facebook, bearing in mind that it changed its name in 2021: https://about.fb.com/news/2021/10/facebook-company-is-now-meta/).

Then, when you are done standardizing `payee_clean`, export it as a new CSV file to your data folder, *giving that file a different name* and read it back into RStudio as a new dataframe.

Then, using that `payee_clean` column, write code to find all of the spending where Facebook is the payee and calculate the total amount of spending for each committee and present the results showing the committee that spent the most first. Which committee is that?

**A1.**
Franchot Peter Friends Of (Friends of Peter Franchot)
```{r}
media_expenses <- maryland_expenses %>% 
  filter(!payee_name == "") %>% 
  filter(str_to_lower(expense_category) == "media")

write_csv(media_expenses, "data/media_expenses.csv")

media_clean <- read.csv("data/media_expenses_clean.csv")

facebook <- media_clean %>% 
  filter(payee_clean == "Facebook/Meta") %>% 
  group_by(committee_name) %>% 
  summarize(
    tot_spending = sum(amount)
  ) %>% 
  arrange(desc(tot_spending))

```

**Q2.** Let's do some research on the top committees from question 1. Look up the top 10 spenders: if they are candidate committees, what offices were they running for? If they were not candidates, what issues do they seem to care about? Write some sentences answering those questions, then go the Facebook ad library (https://www.facebook.com/ads/library/?active_status=all&ad_type=political_and_issue_ads&country=US&sort_data\[direction\]=desc&sort_data\[mode\]=relevancy_monthly_grouped&media_type=all) and look for ads from the top spender. What do you notice about those ads?

**A2.** 
The top two spenders were running for governor and lieutenant governor together. Next was a PAC that supported progressive candidates who support immigrant & working families. Fourth was another candidate for governor. Fifth was another PAC that supported candidates who did things for working families and for "the many." The sixth was a candidate running for Harford County Executive. After that was another governor and lieutenant governor running pair. Then another democratic primary governor candidate. After that was a candidate running for a county council seat in Anne Arundel county. Finally, the tenth highest spender was a republican representing Maryland District 6.

The top spender, Peter Franchot, had advertising done in 2021. Any ads placed in 2022 were done by the combination slate of him and his running partner. 

**Q3.** Given the list of top spenders in A1 and the primary results for governor (https://elections.maryland.gov/elections/2022/primary_results/gen_results_2022_1.html), who is missing from that A1 list? Did those campaigns run Facebook ads, according to Facebook's ad library? Who are the biggest recipients of media expenses by the Democratic primary winner for governor (HINT: that person has two committees)?

**A3.** 
The most glaring campaigns that are missing are those of Wes Moore and Aruna Miller, Dan Cox and his running partner, and Tom Perez and his running mate.

Wes Moore did run facebook ads, as did Dan Cox and Tom Perez.

The biggest recipient for Wes Moore by a long shot is SKDK, followed by Foglamp and Apollo Artistry
```{r}
moore_ads <- media_clean %>% 
  filter(str_detect(committee_name, "Moore")) %>% 
  group_by(payee_clean) %>% 
  summarize(
    tot_spend = sum(amount)
  ) %>% 
  arrange(desc(tot_spend))
```

**Q4.** The media expense data you cleaned also has a standardized purpose column. Write code to generate a list of the committees with the total amount of spending on online advertising. How does this list compare to the one from Q1 about top Facebook spenders? Of the top 5 committees who spent the most on Facebook ads from Q1, are there any that appear to have spent *all* of their online advertising dollars on Facebook?

**A4.** 
The combination Franchot Anderson-Walker slate seems to have spent all their online advertising money on Facebook. The same is true of Ashwani Jain's campaign.
```{r}
media_purpose <- media_clean %>% 
  filter(expense_purpose == "Online Advertising") %>% 
  group_by(committee_name) %>% 
  summarize(
    tot_spend = sum(amount)
  ) %>% 
  arrange(desc(tot_spend))
```

**Q5** You've been asked to summarize your findings about spending on Facebook by Maryland political candidates. How much do we know and what don't we know? What's the best story idea you've found here?

**A5**
We know which candidates spent money DIRECTLY on facebook ads, instead of going through a third party like SKDK. If candidates went through third parties, we don't know how much they spent on Facebook advertising. 

I think the best story would be one about how candidates may be getting around showing what they are really spending on by using third party organizations. Or, looking at who actually puts in the work and ideas for campaigns and how candidates perform as a result. AKA do candidates not using a third party firm stand a chance against candidates who do use one?