---
title: "lab_10"
author: "Khushboo Rathore"
date: "2023-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries, plus some of the map-specific ones from the pre_lab. You'll also be fetching Census data in Q3. Read the questions carefully!

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Load libraries
library(tidyverse)
library(tidycensus)
library(janitor)
library(lubridate)
library(tigris)
```

Let's explore Maryland elections, power outages and electric vehicle data and make some maps! 

## Questions 

**Q1.** Maryland voters made history in November by electing the state's first two African-American statewide office holders: Wes Moore as governor and Anthony Brown as attorney general. You want to compare the two's electoral performances. Using the `md_statewide_county.csv` file in the data folder, calculate the total number of votes for both the governor's race and the attorney general's race, then percentages for both Moore and Brown _based on the total of their respective contests_. Then calculate the difference between their percentages.

Get the geometry for Maryland's counties using tigris and, joining that dataframe to your results dataframe, make a county-level map of the differences between Moore and Brown's percentages. What is the story here, and what is the county with the most interesting results?

**A1.** 
The story here is that while most counties had a higher percent voting for Wes Moore than for Brown, Prince George's County and Baltimore City went the opposite direction. I think Baltimore City is the most interesting results. While PG County is in the same boat, it is also surrounded by counties that didnt have a huge difference in the percentage of voted. However, all the counties directly surrounding Baltimore City had at least a 2.5 percentage point difference. This begs the question - what made that vote so different?
```{r}
county_gov <- read_csv("data/md_statewide_county.csv") %>% 
  clean_names() %>% 
  mutate(tot_gov = moore + cox + lashar + wallace + harding + sefcik + gov_writein) %>% 
  mutate(tot_ag = peroutka + brown + ag_writein) %>% 
  select(geoid, county, moore, tot_gov, brown, tot_ag) %>% 
  mutate(percent_moore = moore/tot_gov*100) %>% 
  mutate(percent_brown = brown/tot_ag*100) %>% 
  mutate(mb_dif = percent_moore-percent_brown)

md_counties <- counties() %>%
  filter(STATEFP == "24") %>% 
  clean_names() %>% 
  mutate(namelsad = case_when(
    str_detect(namelsad, "city")~"Baltimore City",
    TRUE~namelsad
  )) %>% 
  rename(county = namelsad)

elect_map <- county_gov %>% 
  mutate(county = case_when(
    str_detect(county, "City|County")~county,
    str_detect(county, "Mary")~"St. Mary's County",
    TRUE~paste(county, "County")
  )) %>%
  mutate(geoid = as.character(geoid)) %>% 
  left_join(md_counties, by = "geoid")

ggplot() +
  geom_sf(data=elect_map, aes(fill=mb_dif, geometry = geometry)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma")

```

**Q2.** Make a county-level map to explore the number of customers impacted by power outages in Maryland. Load the `county_outages_march2023.csv` file in the data folder and calculate a rate of customers with outages per 10,000 total customers for each county. Then join that dataframe to your Maryland counties dataframe you made in Q1 and create a county-level map of the results; you may need to tweak the fill scale to make the distinctions clearer. What's the potential story here? Given access to other months' outage data, what would your next steps be in reporting a story?

**A2.** 
I think that the potential story is about why Talbot County has so many outages. Currently, it looks like areas closer to the ocean have a higher number of outages per 1,000 customers. If I had data from other months, I would first combine the datasets and see an average rate over the course of a year. I would map this and then also compare the data for each month, either in table form or through 12 different maps. I would want to see if there are some times of the year where different areas are impacted more. After comprehensively looking at the data, I might also want to look at a trend over the course of multiple years as the climate has been changing and look at population density changes.
```{r}
county_outages <- read_csv("data/county_outages_march2023.csv") %>% 
  mutate(outage_rate = outage_customers/total_customers*1000)

outage_map <- county_outages %>% 
  mutate(county = case_when(
    str_detect(county, "City|County")~county,
    TRUE~paste(county, "County")
  )) %>%
  left_join(md_counties, by = "county")
  
ggplot() +
  geom_sf(data=outage_map, aes(fill=outage_rate, geometry = geometry)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma", trans = "log")
```

**Q3.** You've been assigned a story about the use of electric vehicles in Maryland and you want to make a map of the prevalence of electric vehicles in each county. Load the `ev_2020_2023.csv` file, which has the number of electric and hybrid vehicles per county from July 2020 and March 2023 and calculate a percentage change for each county (remember, that's `new-old/old`). Using the Maryland counties dataframe you made for Q1, make a county map showing the percentage change.What counties stand out to you? What questions do you have about the changes you're seeing?

Next, use tidycensus to retrieve county population data for Maryland from 2021 (you did this in the pre_lab) and join it to your vehicles dataframe. HINT: you may not need to join it on the county name, and you should already have the geometry!

Then, calculate a per capita rate of electric & hybrid vehicles using the March 2023 figure and the population estimate. You will need to choose the "per capita" value.

Finally, make a second county map showing the per capita rate. What counties stand out on this map? How is it different from the first one you made?

**A3.** 
There was obviously significant positive change in all counties in terms of EV car usage. The county that stands out the most to me is Caroline county, which more than quadrupled the number of electric cars. However, it's interesting that Caroline borders the county with the lowest percent difference, Dorchester county. It makes me wonder if those two are related since the majority of the other values are in the same middle range.

In the second map, Montgomery, Howard and Somerset county stand out the most. Montgomery and Howard don't surprise me since those are wealthier counties and the rate of EV/Hybrids tends to be higher. According to census reporter, Somerset county is among the most impoverished counties in Maryland. This indicates that there may be some connection in poverty and ability to afford electric or hybrid vehicles, which tend to be more expensive. 

Unlike the first map, the second more accurately reflects what the data means. They both show the data, but the first is likely more indicative of population growth/movement versus the second, which may indicate actual problems with the electric vehicle/hybrid industry.

```{r}
ev_data <- read_csv("data/ev_2020_2023.csv") %>% 
  mutate(percent_diff = (march_2023-july_2020)/july_2020*100)

ev_map <- ev_data %>% 
  mutate(county = case_when(
    str_detect(county, "City|County")~county,
    TRUE~paste(county, "County")
  )) %>%
  left_join(md_counties, by = "county")

ggplot() +
  geom_sf(data=ev_map, aes(fill=percent_diff, geometry = geometry)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma", trans = "log")

md_county_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              year = 2021,
              state = "MD") %>% 
  clean_names() %>%
  select(geoid, name, estimate)

car_census <- ev_map %>%
  left_join(md_county_population, by = "geoid") %>% 
  mutate(per_capita =  march_2023/estimate*10000)

ggplot() +
  geom_sf(data=car_census, aes(fill=per_capita, geometry = geometry)) +
  theme_minimal() +
  scale_fill_viridis_b(option="magma", trans = "log")
```

